---
output: github_document
---

# tidychangepoint

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/beanumber/tidychangepoint/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/beanumber/tidychangepoint/actions/workflows/R-CMD-check.yaml)
  <!-- badges: end -->

## Usage

To install `tidychangepoint`:

```{r, eval=FALSE}
remotes::install_github("beanumber/tidychangepoint")
```

To load it:

```{r, message=FALSE}
library(tidychangepoint)
```

## Tidy methods for changepoint analysis

The `tidychangepoint` package allows you to use any number of algorithms for detecting changepoint sets in univariate time series with a common, `tidyverse`-compliant interface. 
Currently, algorithms from `changepoint`, `wbs`, and several genetic algorithms made accessible via `GA` are supported. 
It also provides model-fitting procedures for commonly-used parametric models, tools for computing various penalty functions, and graphical diagnostic displays. 

Changepoint sets are computed using the `segment()` function, which takes a numeric vector that is coercible into a `ts` object, and a string indicating the algorithm you wish you use. 
`segment()` always returns a `tidycpt` object. 

```{r, message=FALSE}
x <- segment(CET, method = "pelt", minseglen = 3)
class(x)
```

Various methods are available for `tidycpt` objects. 
For example, `as.ts()` returns the original data as `ts` object, and `changepoints()` returns the set of changepoint indices. 

```{r}
changepoints(x)
```

If the original time series has time labels, we can also retrieve that information. 

```{r}
changepoints(x, use_labels = TRUE)
```


The `fitness()` function returns the both the value and the name of the objective function that the algorithm used to find the optimal changepoint set. 

```{r}
fitness(x)
```


## Citation

```{r}
citation("tidychangepoint")
```

