---
title: "Climate data in tidychangepoint"
bibliography: refs.bib
---

The `tidychangepoint` package provides three climate-related time series. 

```{r setup}
library(tidychangepoint)
```

## Central England Temperature

@shi2022changepoint use changepoint detection algorithms to analyze a time series of annual temperature data from Central England. 
These data are available via `CET` from `tidychangepoint`. 

These data go back to 1659, and a simple plot illustrates the increase in temperature in recent years. 

```{r}
plot(CET)
```

@shi2022changepoint use a genetic algorithm to identify changepoints in this time series. 
The code below reproduces this analysis. 
Note that since the genetic algorithm is random, results vary. 
Shi, et al. used a `maxiter` value of 50,000 in order to obtain the results used in the paper. 
Here, we use a much lower value solely in the interest of computational speed. 
Note that this algorithm is fitting a "meanshift" model, which estimates the mean $\mu_i$ for each of the regions defined by the changepoint set. 
The objective function employs the `BIC` penalty. 
One departure from Shi's implementation is the use of the `log_gabin_population()` function to generate the first generation of 200 (i.e. `popSize`) possible changepoint sets. 
Each data point is chosen uniformly at random with probability equal to $\ln{N}$, where $N$ is the number of observations (362, in this case). 

```{r ga-shi}
trend_wn <- CET |>
  segment(
    method = "ga", 
    model_fn = fit_meanshift, 
    penalty_fn = BIC, 
    population = log_gabin_population(CET),
    popSize = 200, 
    maxiter = 50,
    run = 10
  )
```

Compare this with the changepoint set discovered by the algorithm:

```{r}
changepoints(trend_wn)
changepoints_labels(trend_wn)
```

The `fitness()` function returns a named vector with the value of the objective function from the discovered changepoint set. 

```{r}
fitness(trend_wn)
```

Information about the regions, including their means, are shown by the `tidy()` function. 

```{r}
tidy(trend_wn)
```

By default, `glance()` returns a summary of the **segmenter** that produced the discovered changepoint set. 
This includes the fitness, the elapsed time, and the parameters used by the segmenter. 

```{r}
glance(trend_wn)
```

However, we can also run `glance()` on the model resulting from the discovered changepoint set. 
This provides information about the model and its fit, including the values of various alternative model fitting metrics.

::: {.callout-important}
It is important to note that only one of these metrics (in this case, BIC) is actually the one used by the segmenter!
:::

```{r}
glance(trend_wn$model)
```

The `plot()` function returns an informative plot of the original time series, with the changepoint set and the corresponding regions demarcated. 

```{r cet_trend_wn}
plot(trend_wn, use_time_index = TRUE)
```


### Comparison to reported values

The changepoint set reported by @shi2022changepoint is $\{1700, 1739, 1988\}$. 
For that configuration with the trendshift model with white noise errors, Table 2 of @shi2022changepoint reports model variance $\hat{\sigma}^2$ of 0.291, a log-likelihood of -290.02, BIC of 650.74, and MDL of 653.07. 

Fitting the trendshift model with white noise errors and running the `glance()` function reveals an exact match to the reported figures. 

```{r}
target_cpts <- c(1700, 1739, 1988)
ids <- time2tau(target_cpts, as_year(time(CET)))
CET |>
  fit_trendshift(tau = ids) |>
  glance()
```

Modifying the model to incorporate AR(1) lagged errors also matches the figures from Table 2. 

```{r}
CET |>
  fit_trendshift_ar1(tau = ids) |>
  glance()
```

## Bogotá particulate matter

```{r}
plot(bogota_pm)
```

## Medellín rainfall

```{r}
plot(mde_rain_monthly)
```

# References
