---
title: "Using Coen's algorithm"
bibliography: refs.bib
---

Please see @baumer2024tidy for more details. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidychangepoint)
```

```{r}
plot(bogota_pm)
```

## Using the original implementation of Coen's algorithm

```{r}
x <- segment(bogota_pm, method = "coen", num_generations = 5)
changepoints(x)
plot(x)
```

## Using the GA implementation of Coen's algorithm

```{r}
y <- segment(bogota_pm, method = "ga-coen", maxiter = 50, run = 10)
changepoints(y)
plot(y)
diagnose(y$model)
tidy(y)
glance(y)
```

### Changing the threshold

By default, the threshold is set to the mean of the observed values, but it can be changed using the `model_fn_args` argument to `segment()`. 

Please note that the number of iterations (`maxiter`) of the genetic algorithm has been set very low here for ease of compilation. 
**NOTA BENE**: To obtain more robust result, set `maxiter` to be something much higher. 
You can also experiment with the `popSize` argument to `segment()`. 

```{r}
z <- segment(
  bogota_pm, 
  method = "ga-coen", 
  maxiter = 5,
  model_fn_args = list(threshold = 50)
)
changepoints(z)
plot(z)
diagnose(z$model)
tidy(z)
glance(z)
```
